# Editing the book

The contents of the book are written in **RMarkdown**. You can use any formatting code that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$. Moreover, you can include chunks of R-code, like this:

<pre><code>```{r}
print("This is an R-command!")
```</code></pre>

The results of these chunks can be rendered to the GitBook:

```{r, echo = FALSE}
print("This is an R-command!")
```

To edit the book, you can change the text in the `.Rmd` files. Each Rmd file should contain **one and only one** chapter. A chapter is defined by the first-level heading `#`, e.g., `# Editing the book`.

Any sub-headings within the chapter are indicated with several `#` signs, e.g., `##` (level 2) and `###` (level 3).

## Creating new chapters

To create a new chapter, you must follow two steps: 1) Create the file, and 2) Include it in the list of chapters.

First, to create the file for a new chapter in Rstudio, click `File > New File > Text file`. At the top of the file, write your chapter heading, as explained above. Then, click `File > Save`. Save the file as `.Rmd`, without spaces in the file name, e.g.: `editing_the_book.Rmd`.

Second, to include it in the list of chapters, open the file `_bookdown.yml` (click it in the Files explorer in the bottom right of Rstudio). This file has a list of `.Rmd` files to be included in the book. In this example, the list looks like this:

```{r, results = "asis"}
tmp <- readLines("_bookdown.yml")
cat(tmp[grep("^rmd_files", tmp):grep("references\\.Rmd", tmp)], sep = "\n")
```

Insert the file name of your new chapter in the desired position in this list.

## Linking across chapters

You can label chapter and section titles using `{#label}` after them. The labels can be used as cross-references. For example, we can link to Chapter \@ref(figtab). If you do not manually label chapters, there will be automatic labels anyway, e.g., Chapter \@ref(examples).

## Advanced editing

The convenient [Rmarkdown Cheat Sheet](https://rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf) by Rstudio covers most of the knowledge required for advanced Rmarkdown editing. You can print it out and stick it to your wall!
